---
title: "Slater is bad at R please change this title"
author: "David Josephs, Andy Heroy, Carson Drake, Che' Cobb"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: united
    highlight: haddock
    df_print: paged
    keep_md: TRUE
    fig_width: 12
    fig_height: 12
    fig_retina: true
---

```{r setup, include = F}
knitr::opts_chunk$set(echo = F, warn = F, message = F, dev = "svg", fig.align = "center")
```

```{r set-options, echo=FALSE, cache=FALSE}
options( digits = 2)
```

# Introduction

TODO: CLEAN UP
Businesses today often need to know where items (such as people or machinery) at any given point in time, in a specified area.  Tracking items indoors provides an interesting challenge as conventional methods (GPS) for establishing location don’t work well indoors.  Nolan and Lang propose an innovative solution to this problem by combining machine learning techniques (K-Nearest Neighbors), and wifi signals in order to create an indoor map that can locate and estimate where a given object/person/thing by assessing its signal strength at various access points (wifi routers) placed throughout that area.  This information proves vital to optimizing workflows for how objects move throughout a space, and how improve upon their future handling to best accommodate the business’s needs.

Initially researchers mapped the static signal strengths of of 7 access points throughout the desired space.  These routers communicate with a scanning device that was methodically placed at known intervals around the desired space.  This collection of data makes up the offline data.  Which can be found at http://rdatasciencecases.org/Data/offline.final.trace.txt.  The raw data is arranged by router, and each of the variables are described below. 

```{r tableguy, fig.cap = "variable description"}
pander::pander(
 list(
      t = "Time stamp (Milliseconds) since 12:00am, January 1, 1970",
      Id = "router MAC address",
      Pos = "Router location",
      Degree = "Direction scanning device was carried by the researcher, measured in Degrees",
      MAC = "MAC address of either the accessrouter, or scanning device combined with corresponding values for signal strength (dBm), the mode in which it was operating(adhoc scanner = 1, access router = 3), and its corresponding channel frequency."
 ,
 Signal = "Received Signal Strength in DbM")
)
```

# Data Cleansing
After removing extraneous comments and formatting the data into a tabular form, we rounded the angles into discrete units of 45 degrees. We also removed data which was not pertinent to the study, such as position in the Z axis, as well as the MAC of the scanner, keeping only access points.  We also removed data from MAC addresses which were similarly not useful to this study. The main MACS we focused on were 5 linksys routers, as well as two additional Alpha routers with the id’s (00:0f:a3:39:e1:c0 and 00:0f:a3:39:dd:cd).  Nolan et al. dropped Alpha router 00:0f:a3:39:dd:cd in their analysis. We will investigate both Alpha router addresses and assess which would be a better use for an RTLS system.

The next step is to create new features focusing on the offline signal strengths of all routers and calculate basic statistics behind each access point.  (median, length, mean, std deviation, etc) with respect to each angle.   
 


```{r processline}
# first we define the processline function, 
# which unsurprisingly processes a single line of the offline or online.txt
library(tidyverse)
processLine = function(x) {
  # here we split the line at the weird markers. Strsplit returns a list
  # we take the first item of the list
  tokens = strsplit(x, "[;=,]")[[1]]
  if (length(tokens) == 10){
    return(NULL)
  }
  # now we are going to stack the tokens
  tmp = matrix(tokens[ - (1:10) ], , 4, byrow = TRUE)
  cbind(matrix(tokens[c(2, 4, 6:8, 10)], nrow(tmp), 6,byrow = TRUE),
  tmp)
}
```

```{r roundOrientation}
roundOrientation = function(angles) {
  refs = seq(0, by = 45, length  = 9)
  q = sapply(angles, function(o) which.min(abs(o - refs)))
  c(refs[1:8], 0)[q]
}
```

```{r readoffline, cache = TRUE}
# this reads in the data
readData <- function(filename, 
                     subMacs = c("00:0f:a3:39:e1:c0", "00:0f:a3:39:dd:cd", "00:14:bf:b1:97:8a",
                                 "00:14:bf:3b:c7:c6", "00:14:bf:b1:97:90", "00:14:bf:b1:97:8d",
                                 "00:14:bf:b1:97:81")) {
  # read it in line by line
  txt = readLines(filename)
  # ignore comments
  lines = txt[ substr(txt, 1, 1) != "#" ]
  # process (tokenize and stack) each line
  tmp = lapply(lines, processLine)
  # rbind each elemnt of the list together
  offline = as.data.frame(do.call(rbind, tmp), 
                          stringsAsFactors= FALSE) 
  # set the names of our matrix
  names(offline) = c("time", "scanMac", 
                     "posX", "posY", "posZ", "orientation", 
                     "mac", "signal", "channel", "type")

  # keep only signals from access points
  offline = offline[ offline$type == "3", ]

  # drop scanMac, posZ, channel, and type - no info in them
  dropVars = c("scanMac", "posZ", "channel", "type")
  offline = offline[ , !( names(offline) %in% dropVars ) ]

  # drop more unwanted access points
  offline = offline[ offline$mac %in% subMacs, ]

  # convert numeric values
  numVars = c("time", "posX", "posY", "orientation", "signal")
  offline[numVars] = lapply(offline[numVars], as.numeric)

  # convert time to POSIX
  offline$rawTime = offline$time
  offline$time = offline$time/1000
  class(offline$time) = c("POSIXt", "POSIXct")

  # round orientations to nearest 45
  offline$angle = roundOrientation(offline$orientation)

  return(offline)
}

offline_original <- readData("../Data/offline.final.trace.txt")
```

# Analysis

## Comparison of Signal strength to angle

For the first stage of our analysis, we created a boxplot at a fixed position of the signal strength vs discrete angle of the two routers in question:

```{r firstplot, fig.cap = "Note how the signal is much weaker at the MAC address on the left compared to the MAC address on the right"}
bad_mac <-  "00:0f:a3:39:dd:cd"
good_mac <- "00:0f:a3:39:e1:c0"
fixfonts <- theme(text = element_text(family = "serif", , face = "bold"))
plt_theme <- ggthemes::theme_hc()  + fixfonts

offline_original %>% mutate(angle = factor(angle)) %>%
  filter(posX == 2 & posY == 12 & mac %in% c(good_mac, bad_mac)) %>%
  ggplot() + geom_boxplot(aes(y = signal, x= angle)) + 
  facet_wrap(. ~ mac, ncol = 1) +  
  ggtitle("Signal vs Angle for a fixed position at selected MACS") + plt_theme
```


It is apparent from the figure above that the signal is much weaker at the router which Nolan and Lang chose not to address. We can look into this further by creating a table of average signal strengths at all positions for the two routers in question:

```{r routertable, fig.cap = "Average signal strengths at the two routers"}
offline_original %>% mutate(angle = factor(angle)) %>%filter(mac %in% c(good_mac, bad_mac)) %>%group_by(mac) %>% summarise(signal_avg = mean(signal), signal_std = sd(signal), signal_iqr = IQR(signal))
# # A tibble: 2 x 4
#   mac               signal_avg signal_std signal_iqr
#   <chr>                  <dbl>      <dbl>      <dbl>
# 1 00:0f:a3:39:dd:cd      -70.5       8.13         13
# 2 00:0f:a3:39:e1:c0      -53.7       5.80          8
```

From this, we can see that the signal at `00:0f:a3:39:dd:cd` is much weaker than the signal at the other router. It also has a much larger standard deviation, and a larger inter quartile range. This could be indicative of why Nolan and Lang did not choose to use this router.

## Signal Strength Density

We also wanted to analyze the signal strength density at each angle of the two routers. This would allow us to not only see signal strenght, but potentially signal clarity, or the separation of the signals at each angle.

```{r dens1, fig.cap = "The signals are much stronger and more separated by angle in the second router"}
offline_original %>% mutate(angle = factor(angle)) %>%
  filter(posX == 2 & posY == 12 & mac %in%  c(good_mac,bad_mac)) %>%
  ggplot(aes(signal, fill = angle )) + geom_density()+
  facet_wrap(. ~ mac, ncol = 1) +  
  ggtitle("Per Angle Signal Density at the Two MACS") + plt_theme + scale_fill_viridis_d()
offline_original$posXY = paste(offline_original$posX, offline_original$posY, sep = "-")
```


From this graph, we can see that the signal is not only stronger at the router located at (00:0f:a3:39:e1:c0) , but also that it has a different distribution at each angle, which would make it easier to identify position. This also likely led to Nolan and Lang’s choice of dropping the router at 00:0f:a3:39:dd:cd.  

## TODO: SOMETHING SOMETHING DISTANCE BULLSHIT

## KNN WHATEVER

To determine practically whether or not the MAC at address `00:0f:a3:39:dd:cd` is better or worse than the MAC at address `00:0f:a3:39:e1:c0`, we ran a K nearest neighbors prediction experiment. The experimental procedure is as follows. First, we create three groups of data, one group which has all MACs except for  `00:0f:a3:39:e1:c0`, one group which has all MACs except for `00:0f:a3:39:dd:cd`, and one group which contains all the data from all the MACs. Then, we tuned to find the optimal value of K using cross validation at a single random seed. Next, we used something something ANOVA to find the best set!!

First, we filtered in the data, selecting only information which was pertinent to our K nearest neighbors analysis. 

```{r}
online <- readData("../Data/online.final.trace.txt")
online$posXY = paste(online$posX, online$posY, sep = "-")

datasets <- list()
datasets$offline_good_mac <- offline_original %>% filter(mac != bad_mac) %>% select(posXY, posX, posY, mac, orientation,angle, signal)
datasets$online_good_mac <- online %>% filter(mac != bad_mac)%>% select(posXY, posX, posY, mac,orientation, angle, signal)
datasets$online_bad_mac <- online %>% filter(mac != good_mac)%>% select(posXY, posX, posY, mac,orientation, angle, signal)
datasets$offline_bad_mac <- offline_original %>% filter(mac != good_mac)%>% select(posXY, posX, posY, mac,orientation, angle, signal)
datasets$offline_all_mac <- offline_original %>% select(posXY, posX, posY, mac,orientation, angle, signal)
datasets$online_all_mac <- online %>% select(posXY, posX, posY, mac, angle, signal)

off_angs <- offline_original %>% arrange(angle) %>% .$angle %>% as.numeric %>% unique
on_angs <- online %>% arrange(angle) %>% .$angle %>% as.numeric %>% unique
# should be all zero, to verify we are rounding the angles
# off_angs - on_angs
```

Next, we spread out the signal data by mac, averaging where there were repeats:
TODO REPHRASE

```{r, echo = TRUE}
prepare_data <- function(df) {
  df %>% pivot_wider(names_from = mac, values_from = signal, values_fn = mean) %>% na.omit
}

datasets_wide <- lapply(datasets, prepare_data)

datasets_wide[[1]] %>% data.frame
datasets_wide[[2]] %>% data.frame
```

Next, we wrote an R S3 class and extended R's `predict` method to include a fast, vectorized KNN, from [here]( http://enhancedatascience.com/2018/05/23/create-your-machine-learning-library-from-scratch-with-r-3-5-knn/):

```{r, echo=T}

# set the stage for storing neighbors etc
KNN <- function(x, y, k = 5){
  if (!is.matrix(x)) {
    x  <-  as.matrix(x)
  }
  if (!is.matrix(y)) {
    y  <-  as.matrix(y)
  }
  results <- list()
  results$points <- x
  results$value  <- y
  results$k  <- k
  results <- structure(results, class = "KNN")
  return(results)
}

# f ∥xi−xj∥2=∥xi∥2+∥xj∥2−2(xi⋅xj)
# https://www.r-bloggers.com/pairwise-distances-in-r/
# https://blog.smola.org/post/969195661/in-praise-of-the-second-binomial-formula
compute_pairwise_distance=function(X,Y){
  xn = rowSums(X ** 2)
  yn = rowSums(Y ** 2)
  outer(xn, yn, '+') - 2 * tcrossprod(X, Y)
}

# prediction method
predict.KNN = function(my_knn, newdata, distance = compute_pairwise_distance){
  if (!is.matrix(newdata))
  {
    newdata = as.matrix(newdata)
  }
  ##Compute pairwise distance
  dist_pair = distance(newdata,my_knn$points)
  # rank distances computing a lovely graph
  crossprod(
            apply(dist_pair,1,order) <= my_knn[['k']], 
            my_knn[["value"]]) / my_knn[['k']]  
  # turn points more than k neighbors away  to zero, otherwise one 
  # crossproduct by value
}

# Next we need to write a function which maps positionXY to numbers and numbers to positionXY
pos_map_fct <- as.factor(datasets[[1]]$posXY)
pos_map_int <- as.integer(pos_map_fct)
# some dumb shit wiht numerics
ok <- datasets[[3]]$posXY
ok2 <- as.integer(as.factor(ok))
ok2
```


Here is the MAC in question

```{r secondplot, fig.cap = "Signal vs Orientation of the much hated bad mac"}
offline_original %>% mutate(angle = factor(angle)) %>%
  filter(posX == 2 & posY == 12 & mac == bad_mac) %>%
  ggplot() + geom_boxplot(aes(y = signal, x= angle)) + 
  facet_wrap(. ~ mac)  + 
  ggtitle("Signal vs Orientation at the missing MAC") + plt_theme
```


Comment on differences fix captions, this mac looks DAMN GOOD TO ME


Recreate that stupid plot, and do it for the other MAC, then improve


```{r dens0, fig.cap = "dumb"}
offline_original %>% mutate(angle = factor(angle)) %>%
  filter(posX == 2 & posY == 12 & mac != bad_mac) %>%
  ggplot(aes(signal)) + geom_density()+
  facet_wrap(mac ~ angle) +  
  ggtitle("Signal Density at each angle of each Mac Except the One we hate") + plt_theme + ggthemes::scale_fill_hc()
```


```{r hated00}
offline_original %>% mutate(angle = factor(angle)) %>%
  filter(posX == 2 & posY == 12 & mac == bad_mac) %>%
  ggplot(aes(signal)) + geom_density()+
  facet_wrap(mac ~ angle) +  
  ggtitle("Signal Density Per angle at the hated MAC") + plt_theme + ggthemes::scale_fill_hc()
```


```{r hated01, fig.cap = "dumb"}
offline_original %>% mutate(angle = factor(angle)) %>%
  filter(posX == 2 & posY == 12 & mac == bad_mac) %>%
  ggplot(aes(signal, fill = angle )) + geom_density()+
  facet_wrap(. ~ mac) +
  ggtitle("Signal density per angle at the hated mac") + plt_theme + scale_fill_viridis_d()
```


# code stops at ipynb chunk 39

We need to do the dumb chunks from 39 to 47, I got the rest

Read in the onlineSummary 

Figure out fig sizes pls

# Appendix

```{r getlabels}
labs = knitr::all_labels()
labs = labs[!labs %in% c("setup", "toc", "getlabels", "allcode")]
```

```{r allcode, ref.label = labs, eval = FALSE, echo = TRUE}
```

#
