---
title: "Team ACDC Case Study 6: Spam Ham, Problem 19"
author: "David Josephs, Andy Heroy, Carson Drake, Che' Cobb"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: readable
    highlight: haddock
    code_folding: hide
    df_print: paged
    keep_md: TRUE
    fig_width: 10
    fig_height: 10
    fig_retina: true
    fig_caption: true
---


```{r setup, include = F}
knitr::opts_chunk$set(warning = F, message = F, dev = "svg", fig.align = "center")
knitr::opts_chunk$set(tidy = T)
knitr::opts_chunk$set(comment = '#>')
options(digits = 2)

```

# Introduction

Spam email is an ever present annoyance in most of modern day life. This case study is to examine the Spam Assassin dataset (http://spamassassin.apache.org) in which we evaluate 9000 emails with spam filters by both classification tree's and recursive partitioning.  To do this, the libararies rpart and caret are utilized and we attempt to improve upon model classification accuracies by exploring the many parameters available in the rpart.control function.  

#TODO - expand intro.  Kinda sucks 

we've selected question 19 from the Nolan and Lang book for our analysis and exploration. The question states. 

Consider the other parameters that can be used to control the recursive partitioning process. Read the documentation for them in the rpart.control() documentation. Also, carry out an Internet search for more information on how to tweak the rpart() tuning parameters. Experiment with values for these parameters. Do the trees that result make sense with your understanding of how the parameters are used? Can you improve the prediction using them?

# Background

The spam assassin dataset is contrived of 9000 emails each comprised of a header, message body and attachment.  The first goal is to understand the design of the email structure in order to correctly indentify what portion of the email is the header, message body and attachments.  After dropping attachments, the headers and message bodies are processed for characteristics that possibly could mark an email as spam.  Looking for characteristics such as capitolization, punctuation or word counts of known spam words that can be found within.  

## Data Description

# TODO - 

# Note.  Could just make a giant r_script with all the functions and load it into here.



```{r dataload}
# we are starting from here, 
# the code used to create the RDA will be shown in the appendix, 
# it seems to run fine on windows but crash on linux
# so this was a good compromise without digging too deeply into the web of 
# lapplys
load("./Data/data.Rda")
emails <- emailDFrp
# probably do the n
# This function...  is strange.
setupRnum = function(data) {
  logicalVars = which(sapply(data, is.logical))
  facVars = lapply(data[ , logicalVars], 
                   function(x) {
                      x = as.numeric(x)
                   })
  cbind(facVars, data[ , - logicalVars])
}

emails = setupRnum(emails)
dim(emails)
emails[is.na(emails)]<-0
df_process <- emails

```

# Analysis

## Minimum Split

```{r}
library(rpart)
library(caret)
library(dplyr)
library(rpart.plot)
set.seed(54)

#Get a count for spam and ham
is_Spam_ct <- table(df_process$isSpam)["T"]
is_Ham_ct <- table(df_process$isSpam)["F"]

#Sample the indexes
testSpamidx = sample(is_Spam_ct, size = floor(is_Spam_ct/3))
testHamidx = sample(is_Ham_ct, size = floor(is_Ham_ct/3))

#Create DF for test and train
df_test_rpart = rbind(df_process[df_process$isSpam == "T",][testSpamidx,],
                      df_process[df_process$isSpam == "F",][testHamidx, ])

df_train_rpart = rbind(df_process[df_process$isSpam == "T",][-testSpamidx,],
                      df_process[df_process$isSpam == "F",][-testHamidx, ])

dim(df_train_rpart)

#Fitting the regression tree.
rpart_fit = rpart(isSpam ~ ., 
                  data = df_train_rpart,
                  minsplit = 2,
                  method = "class")

#Plot it with either prp or rpart.plot
#prp(rpart_fit, extra = 106)
rpart.plot(rpart_fit)

#Now lets make some predictions.  We'll predict on the df_test_rpart, with the fit we just achieved directly above. 

rp_predict <- predict(rpart_fit,
                      newdata = df_test_rpart[, names(df_test_rpart) != "isSpam"],
                      type = "class")

#summary(rp_predict)

spam_predict = rp_predict[ df_test_rpart$isSpam == "T"]
ham_predict = rp_predict[ df_test_rpart$isSpam == "F"]




```
Minsplit vs ntrees and accuracy/related metrics

## Ratio of minsplit and minbuxkets
normally its 3:1, see what happens

## Complexity and interaction with minsplit

### Lots o trees

### Few o trees

## Maxdepth

### Deep and wide

### Deep and narrow

### Short and wide

### Short and narrow

## Surragets
maybe
##xval
maybe

# Conclusion

Use cases for interesitng chouces
