---
title: "Team ACDC Case Study 2: Cherry Blossom"
author: "David Josephs, Andy Heroy, Carson Drake, Che' Cobb"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: united
    highlight: haddock
    code_folding: hide
    df_print: paged
    keep_md: TRUE
    fig_width: 10
    fig_height: 10
    fig_retina: true
---

```{r setup, include = F}
knitr::opts_chunk$set(warning = F, message = F, dev = "svg", fig.align = "center")
knitr::opts_chunk$set(tidy = T)
knitr::opts_chunk$set(comment = '#>')
options(digits = 2)

```

# Introduction
```{r}
library(rvest)
ubase <- 'http://cherryblossom.org/'
url <- paste0(ubase, 'results/2012/2012cucb10m-m.htm')
doc <- read_html(url)
```

```{r}
library(stringr)
extract_res_table <- function(url) {
  read_html(url) %>% 
    html_nodes('pre') %>% 
    html_text() %>% 
    str_split('\\r\\n') %>% 
    .[[1]]
}
```

```{r}
men_urls <- c(
  'results/1999/cb99m.html',
  'results/2000/Cb003m.htm',
  'results/2001/oof_m.html',
  'results/2002/oofm.htm',
  'results/2003/CB03-M.HTM',
  'results/2004/men.htm',
  'results/2005/CB05-M.htm',
  'results/2006/men.htm',
  'results/2007/men.htm',
  'results/2008/men.htm',
  'results/2009/09cucb-M.htm',
  'results/2010/2010cucb10m-m.htm',
  'results/2011/2011cucb10m-m.htm',
  'results/2012/2012cucb10m-m.htm'
)
```


```{r}
men_urls <- paste0(ubase, men_urls)
head(men_urls, 3)
```


```{r}
library(purrr)
men_tables <- map(men_urls, extract_res_table)
```

    


```{r}
map_int(men_tables, length)
```

```{r}
str_sub(men_tables[[1]], 1, 200)
```

```{r}
men_tables[[1]] <- str_split(men_tables[[1]], '\\n')[[1]]
map_int(men_tables, length)
```



```{r}
extract_res_table <- function(url, year = 2001) {
  selector <- if (year == 2000) 'font' else 'pre'
  regexp <- if (year == 1999) '\\n' else '\\r\\n'
  
  result <- read_html(url) %>% 
    html_nodes(selector)
  
  if (year == 2000) result <- result[[4]]
  
  result <- result %>% 
    html_text()
  
  if (year == 2009) return(result)
  
  result %>% 
    str_split(regexp) %>% 
    .[[1]]
}
```


```{r}
years <- 1999:2012
men_tables <- map2(men_urls, years, extract_res_table)
names(men_tables) <- years
map_int(men_tables, length)
```


```{r}
women_urls <- c(
  'results/1999/cb99f.html',
  'results/2000/Cb003f.htm',
  'results/2001/oof_f.html',
  'results/2002/ooff.htm',
  'results/2003/CB03-F.HTM',
  'results/2004/women.htm',
  'results/2005/CB05-F.htm',
  'results/2006/women.htm',
  'results/2007/women.htm',
  'results/2008/women.htm',
  'results/2009/09cucb-F.htm',
  'results/2010/2010cucb10m-f.htm',
  'results/2011/2011cucb10m-f.htm',
  'results/2012/2012cucb10m-f.htm'
)
```


```{r}
women_urls <- paste0(ubase, women_urls)
extract_res_table <- function(url, year = 2001, female = TRUE) {
  selector <- if (year == 2000) 'font' else 'pre'
  regexp <- if (year == 1999) '\\n' else '\\r\\n'
  
  result <- read_html(url) %>% 
    html_nodes(selector)
  
  if (year == 2000) result <- result[[4]]
  
  result <- result %>% 
    html_text()
  
  if (year == 2009 && female == FALSE) return(result)
  
  result %>% 
    str_split(regexp) %>% 
    .[[1]]
}
```


```{r}

men_tables <- map2(men_urls, years, extract_res_table, female = FALSE)
women_tables <- map2(women_urls, years, extract_res_table, female = TRUE)
names(men_tables) <- years
names(women_tables) <- years
map_int(men_tables, length)
map_int(women_tables, length)
```


```{r}
dir.create('men')
dir.create('women')

walk2(men_tables,
      paste('men', paste(years, 'txt', sep = '.'), sep = '/'),
      writeLines)
walk2(women_tables,
      paste('women', paste(years, 'txt', sep = '.'), sep = '/'),
      writeLines)
```

```{r}
findColLocs = function(spacerRow) {
  
  spaceLocs = gregexpr(" ", spacerRow)[[1]]
  rowLength = nchar(spacerRow)
  
  if (substring(spacerRow, rowLength, rowLength) != " ")
    return( c(0, spaceLocs, rowLength + 1))
  else return(c(0, spaceLocs))
}


selectCols = function(shortColNames, headerRow, searchLocs) {
  sapply(shortColNames, function(shortName, headerRow, searchLocs){
    
    startPos = regexpr(shortName, headerRow)[[1]]
    
    if (startPos == -1) return( c(NA, NA) )
    
    index = sum(startPos >= searchLocs)
    c(searchLocs[index] + 1, searchLocs[index + 1])
  }, 
  
  headerRow = headerRow, searchLocs = searchLocs )
}
```


```{r}
extractVariables = 
  function(file, varNames =c("name", "home", "ag", "gun",
                             "net", "time"))
  {
    eqIndex = grep("^===", file)

    spacerRow = file[eqIndex] 
    headerRow = tolower(file[ eqIndex - 1 ])
    body = file[ -(1 : eqIndex) ]

    footnotes = grep("^[[:blank:]]*(\\*|\\#)", body)
    if ( length(footnotes) > 0 ) body = body[ -footnotes ]
    blanks = grep("^[[:blank:]]*$", body)
    if (length(blanks) > 0 ) body = body[ -blanks ]
    
    searchLocs = findColLocs(spacerRow)
    locCols = selectCols(varNames, headerRow, searchLocs)
    
    Values = mapply(substr, list(body), start = locCols[1, ], 
                    stop = locCols[2, ])
    colnames(Values) = varNames
    
    return(Values)
  }
```


```{r}
wfilenames <- list.files('women', pattern = '.txt$', full.names = TRUE)
women_files <- map(wfilenames, readLines)
names(women_files) <- str_match(wfilenames, 'women/(.*).txt')[ ,2]
```


```{r}
mfilenames <- list.files('men', pattern = '.txt$', full.names = TRUE)
men_files <- map(mfilenames, readLines)
names(men_files) <- str_match(mfilenames, 'men/(.*).txt')[ ,2]
```


```{r}
men_res_mat <- map(men_files, extractVariables)
length(men_res_mat)
map_int(men_res_mat, nrow)
```


```{r}
men_file_2001 <- men_files$`2001`
women_file_2001 <- women_files$`2001`

eq_idx_2001 <- str_which(men_file_2001, '^===')
spacer_row_2001 <- men_file_2001[eq_idx_2001]
header_row_2001 <- men_file_2001[eq_idx_2001 - 1] %>% str_to_lower()

women_files$`2001`[2] <- header_row_2001
women_files$`2001`[3] <- spacer_row_2001
```


```{r}
women_res_mat <- map(women_files, extractVariables)
length(women_res_mat)
map_int(women_res_mat, nrow)
```


```{r}
age <- as.numeric(men_res_mat$`2012`[ ,'ag'])
```


```{r}
tail(age)
```


```{r}
age <- map(men_res_mat, ~ as.numeric(.x[ ,'ag']))
```


```{r}
library(tibble)
library(tidyr)
library(ggplot2)

age %>% 
  enframe(name = 'year', value = 'age') %>% 
  unnest() %>% 
  ggplot(aes(year, age)) +
  geom_boxplot()
```



```{r}
sapply(women_res_mat, nrow)
```


```{r}
age <- map(women_res_mat, ~ as.numeric(.x[ ,'ag']))
```


```{r}
age %>% 
  enframe(name = 'year', value = 'age') %>% 
  unnest() %>% 
  ggplot(aes(year, age)) +
  geom_boxplot()
```


```{r}
sapply(age, function(x) sum(is.na(x)))


```{r}
convert_time <- function(t) {
  time_pieces <- str_split(t, ':')
  map_dbl(time_pieces, function(x) {
    x <- as.numeric(x)
    if (length(x) == 2) x[1] + x[2] / 60
    else 60 * x[1] + x[2] + x[3] / 60
  })
}
```


```{r}
create_df = function(Res, year, sex) 
{
  if ( !is.na(Res[1, 'net']) ) useTime = Res[ , 'net']
  else if ( !is.na(Res[1, 'gun']) ) useTime = Res[ , 'gun']
  else useTime = Res[ , 'time']
  
  useTime = gsub("[#\\*[:blank:]]", "", useTime)
  runTime = convert_time(useTime[ useTime != "" ])
  
  Res = Res[ useTime != "", ]
  age = gsub("X{2}\\s{1}?|\\s{3}?","0  ", Res[,'ag'])
  Res[, 'ag'] = age
  
  Results = data.frame(year = rep(year, nrow(Res)),
                       sex = rep(sex, nrow(Res)),
                       name = Res[ , 'name'], home = Res[ , 'home'],
                       age = as.numeric(Res[, 'ag']), 
                       runTime = runTime,
                       stringsAsFactors = FALSE)
  invisible(Results)
}
```


```{r}
separatorIdx = grep("^===", men_files[["2006"]])
separatorRow = men_files[['2006']][separatorIdx]
separatorRowX = paste(substring(separatorRow, 1, 63), " ", 
                      substring(separatorRow, 65, nchar(separatorRow)), 
                      sep = "")
men_files[['2006']][separatorIdx] = separatorRowX

menResMat = sapply(men_files, extractVariables)
menDF = mapply(create_df, menResMat, year = 1999:2012,
               sex = rep("M", 14), SIMPLIFY = FALSE)

separatorIdx = grep("^===", women_files[["2006"]])
separatorRow = women_files[['2006']][separatorIdx]
separatorRowX = paste(substring(separatorRow, 1, 63), " ", 
                      substring(separatorRow, 65, nchar(separatorRow)), 
                      sep = "")
women_files[['2006']][separatorIdx] = separatorRowX

women_files[[3]] = append(women_files[[3]], men_files[[3]][4:5], after=3)

womenResMat = sapply(women_files, extractVariables)

womenDF = mapply(create_df, womenResMat, year = 1999:2012,
                 sex = rep("W", 14), SIMPLIFY = FALSE)
```

```{r}
allMen = do.call(rbind, menDF)
allWomen = do.call(rbind, womenDF)
```


```{r}
head(allWomen)
```



```{r}

```

