---
title: 'Team CCAD Case Study 8: Crox stock ARIMA analysis'
author: "David Josephs, Andy Heroy, Carson Drake, Che' Cobb"
date: '`r Sys.Date()`'
output:
  html_document:
    code_folding: hide
    df_print: paged
    fig_caption: true
    fig_height: 10
    fig_retina: yes
    fig_width: 10
    highlight: haddock
    keep_md: yes
    number_sections: yes
    theme: readable
    toc: yes
  pdf_document:
    toc: yes
---


```{r setup, include = F}
knitr::opts_chunk$set(warning = F, message = F, dev = "svg", fig.align = "center")
#knitr::opts_chunk$set(tidy = T)
knitr::opts_chunk$set(comment = '#>')
options(digits = 2)

```

```{r Load_Libraries_Counter}
library(quantmod)
library(tswge)
library(dplyr)
library(magrittr)
library(tidyverse)

counter <- function() {
  x <- 0
  return (
          function() {
            x <<- x+1
            return(x)
          }

  )
}

cnt <- counter()

cap <- function(str) {
  paste("**Figure", cnt(), ":**",str)
}
```

# Introduction
The purpose of this case study is to fit an ARIMA time series model to a stock of our choosing.   Since we enjoy the finer consumer products avaiable on the market, we chose to write our case study about CROX stock.  Yes, that long forgotten footwear loved by all those who spend too much time on their feet in a given day.  To accomplish this task, we've chosen to utilize the library (TSWGE) which was written and composed by Dr. Woodward, Gray, and Elliot, along the whamo brilliance of Dr. Bivin Sadler.  This library contains many useful functions for time series analysis which we describe and explore in the analysis below.


# Background
To begin, we'll explain what an ARIMA model is comprised of, and how they are useful in time series forecasting techniques.  

# TSWGE

# ARIMA models

# Stock Analysis


```{r getStockData}

crox <- getSymbols("CROX",src = "yahoo")
cstock <- CROX
cstock_2years <-cstock['2018-02-01::2020-02-01']
candleChart(cstock_2years, theme = "white")
str(cstock_2years)
cstock_2years %<>% data.frame

train_indices <- 1:(nrow(cstock_2years) - 5)
crox_train <- cstock_2years[train_indices,]
crox_test <- cstock_2years[-train_indices,]


cadj  <- crox_train$CROX.Close
diff1  <- . %>% artrans.wge(phi.tr = 1)
cadj_diff <- diff1(cadj)
aic5.wge(cadj_diff, type="bic")

```


```{r Anaki_ARIMA_110}
croxd1.est.AR1 <- est.arma.wge(cadj_diff, p = 1, q = 0)
croxd1.fore.AR1 <- fore.aruma.wge(crox_train$CROX.Close, 
                              d = 1,
                              phi= croxd1.est.AR1$phi,
                              n.ahead = 5,
                              limits = T,
                              lastn = F)

plotts.sample.wge(croxd1.est.AR1$res, arlimits = T)
acf(croxd1.est.AR1$res, lag.max = 50)
pacf(croxd1.est.AR1$res)

P24 = ljung.wge(croxd1.est.AR1$res)
P48 = ljung.wge(croxd1.est.AR1$res, K=48)

P24$pval
P48$pval


ARIMA110_ASE = mean((cstock_2years$CROX.Close[(length(cstock_2years$CROX.Close)-4):length(cstock_2years$CROX.Close)] - croxd1.fore.AR1$f)^2)


ARIMA110_ASE
```

```{r Anaki_ARIMA_012}

#Tried fitting with an ARIMA_012.  Keep getting an error though.
croxd1.est.ARIMA.012 <- est.arma.wge(cadj_diff, p = 0, q = 2)
croxd1.for.ARIMA.012 <- fore.aruma.wge(crox_train$CROX.Close, 
                              d = 1,
                              theta = croxd1.est.ARIMA.012$theta,
                              n.ahead = 5,
                              limits = T,
                              lastn = F)

plotts.sample.wge(croxd1.est.ARIMA.012$res, arlimits = T)
acf(croxd1.est.ARIMA.012$res, lag.max = 50)
pacf(croxd1.est.ARIMA.012$res)

P24 = ljung.wge(croxd1.est.ARIMA.012$res)
P48 = ljung.wge(croxd1.est.ARIMA.012$res, K=48)

P24$pval
P48$pval


ARIMA012_ASE = mean((cstock_2years$CROX.Close[(length(cstock_2years$CROX.Close)-4):length(cstock_2years$CROX.Close)] - croxd1.for.ARIMA.012$f)^2)


ARIMA012_ASE
```


```{r}


```

# Grid Search?  
### I think he wants us to run a grid search of p,d,q.  But we'll use AIC5 for that. 


# Conclusion
